\chapter{Further Information}

If you have questions that are not answered by this manual, the
sources of information listed below may be of interest to you.  Note
that bug reports, suggestions and contributions related to the
software (or the documentation) should be sent to the Xen developers'
mailing list (address below).


\section{Other documentation}

If you are mainly interested in using (rather than developing for)
Xen, the \emph{Xen Users' Manual} is distributed in the {\tt docs/}
directory of the Xen source distribution.

% Various HOWTOs are also available in {\tt docs/HOWTOS}.


\section{Online references}

The official Xen web site is found at:
\begin{quote}
{\tt http://www.cl.cam.ac.uk/Research/SRG/netos/xen/}
\end{quote}

This contains links to the latest versions of all on-line
documentation.


\section{Mailing lists}

There are currently four official Xen mailing lists:

\begin{description}
\item[xen-devel@lists.xensource.com] Used for development
  discussions and bug reports.  Subscribe at: \\
  {\small {\tt http://lists.xensource.com/xen-devel}}
\item[xen-users@lists.xensource.com] Used for installation and usage
  discussions and requests for help.  Subscribe at: \\
  {\small {\tt http://lists.xensource.com/xen-users}}
\item[xen-announce@lists.xensource.com] Used for announcements only.
  Subscribe at: \\
  {\small {\tt http://lists.xensource.com/xen-announce}}
\item[xen-changelog@lists.xensource.com] Changelog feed
  from the unstable and 2.0 trees - developer oriented.  Subscribe at: \\
  {\small {\tt http://lists.xensource.com/xen-changelog}}
\end{description}

Of these, xen-devel is the most active.
