\chapter{Further Support}

If you have questions that are not answered by this manual, the
sources of information listed below may be of interest to you.  Note
that bug reports, suggestions and contributions related to the
software (or the documentation) should be sent to the Xen developers'
mailing list (address below).


\section{Other Documentation}

For developers interested in porting operating systems to Xen, the
\emph{Xen Interface Manual} is distributed in the \path{docs/}
directory of the Xen source distribution.


\section{Online References}

The official Xen web site is found at:
\begin{quote} {\tt http://www.cl.cam.ac.uk/netos/xen/}
\end{quote}

This contains links to the latest versions of all online
documentation, including the latest version of the FAQ.

Information regarding Xen is also available at the Xen Wiki at
\begin{quote} {\tt http://wiki.xensource.com/xenwiki/}\end{quote}
The Xen project uses Bugzilla as its bug tracking system. You'll find
the Xen Bugzilla at http://bugzilla.xensource.com/bugzilla/.


\section{Mailing Lists}

There are several mailing lists that are used to discuss Xen related
topics. The most widely relevant are listed below. An official page of
mailing lists and subscription information can be found at \begin{quote}
  {\tt http://lists.xensource.com/} \end{quote}

\begin{description}
\item[xen-devel@lists.xensource.com] Used for development
  discussions and bug reports.  Subscribe at: \\
  {\small {\tt http://lists.xensource.com/xen-devel}}
\item[xen-users@lists.xensource.com] Used for installation and usage
  discussions and requests for help.  Subscribe at: \\
  {\small {\tt http://lists.xensource.com/xen-users}}
\item[xen-announce@lists.xensource.com] Used for announcements only.
  Subscribe at: \\
  {\small {\tt http://lists.xensource.com/xen-announce}}
\item[xen-changelog@lists.xensource.com] Changelog feed
  from the unstable and 2.0 trees - developer oriented.  Subscribe at: \\
  {\small {\tt http://lists.xensource.com/xen-changelog}}
\end{description}
